---
title: "Exploatory Visualization"
author: "Clayton Glasser"
date: "August 21, 2018"
output: html_document
---

## Install Packages

```{r packages}
install.packages('tidyverse')
library(tidyverse)
install.packages('naniar')
library(naniar)
install.packages('ggmap')
library(ggmap)
install.packages('mapproj')
library(mapproj)
```


## Read CSVs into dataframes

```{r }
florence_1 <- read_csv("1424429.csv")
florence_2 <- read_csv("1424435.csv")
northeast <- read_csv("1424479.csv")
southeast <- read_csv("1424485.csv")
```


## Inspect dataframe construct

```{r }
glimpse(florence_1)
glimpse(florence_2)
glimpse(northeast)
glimpse(southeast)

str(florence_1)
str(florence_2)
str(northeast)
str(southeast)

colnames(florence_1)
colnames(florence_2)
colnames(northeast)
colnames(southeast)
```


## Visualize missing data with vis_miss from nanair library. Ignore large data warning.

```{r }
vis_miss(florence_1, warn_large_dat = FALSE) + labs(y = "missing data by variable in florence_1")
vis_miss(florence_2, warn_large_dat = FALSE) + labs(y = "missing data by variable in florence_2")
vis_miss(northeast, warn_large_dat = FALSE) + labs(y = "missing data by variable in northeast")
vis_miss(southeast, warn_large_dat = FALSE) + labs(y = "missing data by variable in southeast")
```


## Create new dataframes that share the same column order (some dfs still missing whole columns)

```{r }
florence_1_reordered <- florence_1 %>% select(STATION, NAME, LATITUDE, LONGITUDE, ELEVATION, DATE, PRCP, PRCP_ATTRIBUTES, SNOW, SNOW_ATTRIBUTES, SNWD, SNWD_ATTRIBUTES, TMAX, TMAX_ATTRIBUTES, TMIN, TMIN_ATTRIBUTES, TOBS, TOBS_ATTRIBUTES)

florence_2_reordered <- florence_2 %>% select(STATION, NAME, LATITUDE, LONGITUDE, ELEVATION, DATE, PRCP, PRCP_ATTRIBUTES, SNOW, SNOW_ATTRIBUTES, SNWD, SNWD_ATTRIBUTES, TAVG, TAVG_ATTRIBUTES, TMAX, TMAX_ATTRIBUTES, TMIN, TMIN_ATTRIBUTES, TOBS, TOBS_ATTRIBUTES)

northeast_reordered <- northeast %>% select(STATION, NAME, LATITUDE, LONGITUDE, ELEVATION, DATE, PRCP, PRCP_ATTRIBUTES, SNOW, SNOW_ATTRIBUTES, SNWD, SNWD_ATTRIBUTES,TAVG, TAVG_ATTRIBUTES, TMAX, TMAX_ATTRIBUTES, TMIN, TMIN_ATTRIBUTES, TOBS, TOBS_ATTRIBUTES)

southeast_reordered <- southeast %>% select(STATION, NAME, LATITUDE, LONGITUDE, ELEVATION, DATE, PRCP, PRCP_ATTRIBUTES, SNOW, SNOW_ATTRIBUTES, SNWD, SNWD_ATTRIBUTES, TAVG, TAVG_ATTRIBUTES, TMAX, TMAX_ATTRIBUTES, TMIN, TMIN_ATTRIBUTES)
```


## Vizualize missing data in the resulting dataframes

```{r }
vis_miss(florence_1_reordered, warn_large_dat = FALSE)
vis_miss(florence_2_reordered, warn_large_dat = FALSE)
vis_miss(northeast_reordered, warn_large_dat = FALSE)
vis_miss(southeast_reordered, warn_large_dat = FALSE)
```

## Unify four dataframes into one by row-binding a on b, forming x, which is bound on c, forming y, which is bound on d, forming z (unified_3). Data is treated as characters to avoid errors. 

```{r }
unified_1 <- bind_rows(mutate_all(florence_1_reordered, as.character), mutate_all(florence_2_reordered, as.character))
unified_2 <- bind_rows(mutate_all(unified_1, as.character), mutate_all(northeast_reordered, as.character))
unified_3 <- bind_rows(mutate_all(unified_2, as.character), mutate_all(southeast_reordered, as.character))
```

## Vizualize the resulting unified dataframe

```{r }
vis_miss(unified_3, warn_large_dat = FALSE)
```

## Write the df to a csv for backup

```{r }
write.csv(unified_3, file = "unified_data.csv",row.names=FALSE, na="")
```

## Convert variables to appropriate object types

```{r }
unified_3$STATION <- as.factor(unified_3$STATION)
unified_3$NAME <- as.factor(unified_3$NAME)
unified_3$LONGITUDE <- as.numeric(as.character(unified_3$LONGITUDE))
unified_3$LATITUDE <- as.numeric(as.character(unified_3$LATITUDE))
unified_3$ELEVATION <- as.numeric(as.character(unified_3$ELEVATION))
unified_3$PRCP <- as.numeric(as.character(unified_3$PRCP))
```

## Vizualization of data missingness within unified_3, grouped by NAME

```{r }
gg_miss_fct(x = unified_3, fct = NAME) + labs(title = "missing data by station in unified_3")

```

## Vizualization of data missingness within unified_3, grouped by variable

```{r }
gg_miss_var(unified_3) + labs(y = "Data missingness in unified_3 grouped by")

```

## Vizualization of data missingness within unified_3, grouped by variable, facted by NAME

```{r }
gg_miss_var(unified_3,
            facet = NAME)

```

## Create a terrain map centered on Mapleton, Oregon big enough to encompass all station locations

```{r }
map <- get_map(
  location = c(-123.858978, 44.031369), # Mapleton Oregon
  source = "google", zoom = 10, maptype = "terrain"
)

```

## Select essential columns for mapping

```{r }
map_data <- unified_3 %>% select(STATION, NAME, LATITUDE, LONGITUDE, ELEVATION, PRCP)

```

## Create a varible containing unified_3 > grouped by STATION >  mutated a "count" column containing the number of observations according to STATION factor > ungroup > retain only unique/distinct rows. I.e. create a dataframe containing one row per STATION and a COUNT column containiing the number of observations in that group. 

```{r }
station_obs <- map_data %>% 
  group_by(STATION) %>% 
  mutate(count = n()) %>% 
  ungroup %>% 
  distinct(STATION, .keep_all = TRUE)

```

## Create a varible station_stats containg the created map, with the station geo-plotted on top, with size according to PRCP data qty and color according to station elevation

```{r }
station_stats <- ggmap(map) + geom_point(data = station_obs, 
                      aes(LONGITUDE, LATITUDE, color=ELEVATION, size=count), alpha=0.75)

```
## Show plot on map with optimized color and size scales

```{r }
print(station_stats +
        scale_colour_gradient(low = "blue", high = "red")) +
        scale_size_continuous(range = c(3, 10))

```
